### 2023-04-10 main/chapter_theory.tex
clarify notation

The Standard Model Lagrangian can be decomposed into this sum of
(partial Lagrangian) terms:
\begin{equation}
\label{eqn:theory_sm_lagrangian_top_sum}
\mathcal{L}_\mathrm{SM} =
\mathcal{L}_\mathrm{YM}
+ \mathcal{L}_\mathrm{WD}
+ \mathcal{L}_\mathrm{Yukawa}
+ \mathcal{L}_\mathrm{Higgs}
,
\end{equation}
which we define below;
$\mathcal{L}_\mathrm{YM}$ corresponds to the gauge bosons,
$\mathcal{L}_\mathrm{WD}$ corresponds to the fermions and their gauge
interactions,
$\mathcal{L}_\mathrm{Yukawa}$ corresponds to the Yukawa interactions between the
fermions and the Higgs,
and
$\mathcal{L}_\mathrm{Higgs}$ corresponds to the Higgs boson and its
potential~\cite{ramond1999journeys,pdg2022ynf}.

@@

\paragraph{Weyl-Dirac.}
The Weyl-Dirac summand $\mathcal{L}_\mathrm{WD}$ implements the fermions and
their gauge interactions, and can be written:
\begin{equation}
\begin{split}
\mathcal{L}_\mathrm{WD} =
\sum_i^3
\Bigl(
& L^\dagger_i \sigma^\mu \mathcal{D}_\mu L_i
+ \bar e^\dagger_i \sigma^\mu \mathcal{D}_\mu \bar e_i \\
& + \mathbf{Q}^\dagger_i \sigma^\mu \mathcal{D}_\mu \mathbf{Q}_i
+ \mathbf{\bar u}^\dagger_i \sigma^\mu \mathcal{D}_\mu \mathbf{\bar u}_i
+ \mathbf{\bar d}^\dagger_i \sigma^\mu \mathcal{D}_\mu \mathbf{\bar d}_i
\Bigr)
,
\end{split}
\end{equation}
where the $\mathcal{D}_\mu$ expressions corresponding to the fermion fields:
$L_i$ for lepton doublets,
$\bar e_i$ for lepton singlets,
$\mathbf{Q}_i$ for quark doublets,
and $\mathbf{\bar u}_i$ and $\mathbf{\bar d}_i$ for up- and down-type quark
singlets, respectively.
The three fermion generations are indexed by $i$.
The quark fields are in colour triplets.

\paragraph{Yukawa.}
The Yukawa term implements masses for the charged leptons by coupling
them to the scalar Higgs field $H$:
\begin{equation}
\mathcal{L}_\mathrm{Yukawa} =
i\hat L_i \bar e_j H^* Y^{[e]}_{ij}
+ i\mathbf{\widehat Q}_i \mathbf{\bar d}_j H^* Y^{[d]}_{ij}
+ i\mathbf{\widehat Q}_i \mathbf{\bar u}_j \tau_2 H Y^{[u]}_{ij}
+ [\textrm{c.c.}]
,
\end{equation}
in which $\tau_2$ contains weak isospin indices,
and $Y^{[e]}_{ij}$, $Y^{[d]}_{ij}$, $Y^{[u]}_{ij}$ are Yukawa coupling matrices
with shape $3\times3$ for the three fermion generations.

\paragraph{Higgs.}
Finally, the Higgs term,
\begin{equation}
\mathcal{L}_\mathrm{Higgs} =
(\mathcal{D}_\mu H)^\dagger (\mathcal{D}^\mu H)
- \left[ -\mu^2 H^\dagger H + \lambda (H^\dagger H)^2 \right]
,
\end{equation}
contains the dynamics and potential for the scalar Higgs doublet.
The potential terms are isolated into square brackets to highlight the negative
sign on the mass term $\mu^2$, which with the quartic term contributes to
spontaneous symmetry breaking~\cite{rubakov1999classical}.

>>>>

The Standard Model Lagrangian (density) can be decomposed into this sum of
(partial Lagrangian) terms:
\begin{equation}
\label{eqn:theory_sm_lagrangian_top_sum}
\mathcal{L}_\mathrm{SM} =
\mathcal{L}_\mathrm{YM}
+ \mathcal{L}_\mathrm{WD}
+ \mathcal{L}_\mathrm{Yukawa}
+ \mathcal{L}_\mathrm{Higgs}
,
\end{equation}
which we define below;
$\mathcal{L}_\mathrm{YM}$ corresponds to the gauge bosons,
$\mathcal{L}_\mathrm{WD}$ corresponds to the fermions and their gauge
interactions,
$\mathcal{L}_\mathrm{Yukawa}$ corresponds to the Yukawa interactions between the
fermions and the Higgs,
and
$\mathcal{L}_\mathrm{Higgs}$ corresponds to the Higgs boson and its
potential~\cite{ramond1999journeys,pdg2022ynf}.
Note that this is only a condensed form of the full Standard Model Lagrangian,
which neglects much of its complexity.

@@

\paragraph{Weyl-Dirac.}
The Weyl-Dirac summand $\mathcal{L}_\mathrm{WD}$ implements the fermions and
their gauge interactions, and can be written:
\begin{equation}
\begin{split}
\mathcal{L}_\mathrm{WD} =
\sum_i^3
\Bigl(
& L^\dagger_i \sigma^\mu \mathcal{D}_\mu L_i
+ \bar e^\dagger_i \sigma^\mu \mathcal{D}_\mu \bar e_i \\
& + \mathbf{Q}^\dagger_i \sigma^\mu \mathcal{D}_\mu \mathbf{Q}_i
+ \mathbf{\bar u}^\dagger_i \sigma^\mu \mathcal{D}_\mu \mathbf{\bar u}_i
+ \mathbf{\bar d}^\dagger_i \sigma^\mu \mathcal{D}_\mu \mathbf{\bar d}_i
\Bigr)
,
\end{split}
\end{equation}
where the $\mathcal{D}_\mu$ expressions corresponding to the fermion fields:
$L_i$ for lepton doublets,
$\bar e_i$ for lepton singlets,
$\mathbf{Q}_i$ for quark doublets,
and $\mathbf{\bar u}_i$ and $\mathbf{\bar d}_i$ for up- and down-type quark
singlets, respectively.
The notation $\bar x$ indicates the Dirac adjoint of $x$.
The three fermion generations are indexed by $i$.
The quark fields are in colour triplets.

\paragraph{Yukawa.}
The Yukawa term implements masses for the charged leptons by coupling
them to the scalar Higgs field $H$:
\begin{equation}
\mathcal{L}_\mathrm{Yukawa} =
i\hat L_i \bar e_j H^* Y^{[e]}_{ij}
+ i\mathbf{\widehat Q}_i \mathbf{\bar d}_j H^* Y^{[d]}_{ij}
+ i\mathbf{\widehat Q}_i \mathbf{\bar u}_j \tau_2 H Y^{[u]}_{ij}
+ [\textrm{c.c.}]
,
\end{equation}
in which $\tau_2$ contains weak isospin indices,
and $Y^{[e]}_{ij}$, $Y^{[d]}_{ij}$, $Y^{[u]}_{ij}$ are Yukawa coupling matrices
with shape $3\times3$ for the three fermion generations.
The notation $H^*$ indicates the complex conjugate of the complex scalar
Higgs field, and `c.c.' implies the complex conjugate of the expression to
its left.

\paragraph{Higgs.}
Finally, the Higgs term,
\begin{equation}
\mathcal{L}_\mathrm{Higgs} =
(\mathcal{D}_\mu H)^\dagger (\mathcal{D}^\mu H)
- \left[ -\mu^2 H^\dagger H + \lambda (H^\dagger H)^2 \right]
,
\end{equation}
contains the dynamics and potential for the scalar Higgs doublet.
The symbol $\mathcal{D}_\mu$ is a gauge-covariant derivative operator.
The notation $x^\dagger$ indicates the hermitian conjugate of $x$.
The potential terms are isolated into square brackets to highlight the negative
sign on the mass term $\mu^2$, which with the quartic term contributes to
spontaneous symmetry breaking~\cite{rubakov1999classical}.
